\section{Evaluation} \label{evaluation}

\subsection{Linear Logic}
While linear logic is more suited to represent real-life situations than classical logic, it showed shortcomings when applied to our domain.
In linear logic all rules are deterministic, meaning every rule has exactly one possible outcome.
There is no possibility for chance.
This forced us to have multiple rules to represent possible different resolutions to some rules.
One example is the rule \lstinline{steal}, which allows an actor to attempt to steal money from another.
There were two possibilities when using this rule: the actor could be successful at stealing or he or she could be caught.
These possibilities should obviously have a different effect on the state of the simulation.
This kind of non-determinism is not supported by linear logic.
To remedy this, we extended our implementation of a linear logic evaluator to support multiple outcomes with authored probability as described in Section \ref{own_linear_logic_evaluator}.

Another drawback of linear logic was that it only knows the resources that exist in the current state.
There is no possibility to test for the absence of a given resource.
While this did not pose a significant problem, it made the rule's setup more complex, both in terms of authorship and in terms of computation.
For many predicates, we now had to introduce an opposite and manually ensure that the resources always appeared mutually exclusive in the resource state.

For example, our system kept track of which characters in the game were related to each other via the predicate \lstinline{related Character Character'}.
When we introduced a new rule \lstinline{get_married}, we wanted to check that the characters getting married were not related to each other.
As linear logic does not allow this, we had to introduce a new predicate \lstinline{not_related} and make sure to assign exactly one of these two predicated during the generation of the initial state.

\subsection{Authorship and Rule Set}
The rules we used in our test runs generated interesting steps of events, but often lacked interest and the agents often fell back to following the same steps of actions to maximize their reward.
We think that given more time, authors with more experience in storytelling would be able to use this kind of system to create a rule set that produces more compelling and varied simulations.
It may be necessary to add more convenience to the pure linear logic system to lower the barrier of entry and allow keeping track of an increasingly complex system of rules.
Further, the resulting outcome may be put into a shape that appears more appealing to players by adding additional flavoring to each rule and dynamically adapting the rendition of sequences of events to the actor, for example by making a more insecure actor's testimony show this character trait when they present the testimony.

When authoring our own rules, we became aware of a responsibility to balance the nature of the rules and the resulting societal norms between what we may think an interesting society should be like and matching what may be the players' expectations, so that they can understand and follow the agent's reasoning.
This includes aspects like the preconditions for marriage, what the consequences for stealing or lying should be like or the rules for when a murder may happen.
This is further amplified by the requirement to attach rewards or penalties to each rule, such that the agents know which actions are desirable and which are better to avoid.

During limited usability tests, players showed great excitement about the often surprising and funny rendition of certain actions by the emojis.
At the same time, certain emotions, such as \enquote{relaxed} and \enquote{fear}, were too ambiguous so that additional explanations were required.
Early on, it became clear that the possession of a knife was too strong of an indicator that someone committed a murder.
When multiple characters with knifes appeared, however, the players were most often able to solve the mystery by investigating their relationship to the victim and reasoning of the possible implications.
As such, the rules we authored appeared to provide strong enough ties to the player's own experiences for them to comprehend the agent's motivation.

\subsection{Flexibility of The System}
To demonstrate the flexibility of our approach, we ported the ruleset to generate simple quest hooks for roleplaying games, such as \enquote{Dungeons and Dragons}.
The outcome is a tool that formulates the flow of events leading up to a moment where the adventurer party should get involved.
The system's output is kept intentionally abstract to let the human author using the tool improvise and fill in details about the story.

Even after investigating different reward models, involving more domain specific characteristics such as physical power or dexterity, we decided to keep using the existing measures, as they proved to be on a level that could easily be mapped to any actor's behavior.
We found the generated stories to encourage thinking outside the box and fueling creativity when it came to developing short plot lines.
However, the range of possible outcomes was of course limited to the ones we authored beforehand, again only providing points of variation in the circumstances that led to the hiring of the story's heroes.
